---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE  # Set to FALSE since we're showing Shiny examples
)
```

# linkeR: Effortless Linked Views for Shiny Applications

<!-- badges: start -->
[![R-CMD-check](https://github.com/JakeWags/linkeR/actions/workflows/R-CMD-check.yaml/badge.svg?cache=bust)](https://github.com/JakeWags/linkeR/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/JakeWags/linkeR/graph/badge.svg)](https://app.codecov.io/gh/JakeWags/linkeR)
[![CRAN status](https://www.r-pkg.org/badges/version/linkeR)](https://CRAN.R-project.org/package=linkeR)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

> **Create synchronized, interactive dashboards where one click updates multiple components**

`linkeR` makes it effortless to create linked views in Shiny applications. When users interact with one component (like clicking a map marker), all related components (tables, charts, other maps) automatically update to show corresponding information.

## What linkeR Does

- **Bidirectional Linking**: Click a map marker → table row highlights. Click table row → map zooms and shows popup.
- **One-Line Setup**: Link multiple components with a single function call
- **Custom Behaviors**: Define exactly what happens when users click different components
- **Multi-Component Support**: Link maps, tables, charts, and more in complex dashboards
- **Zero Boilerplate**: No manual observer setup or event handling required

## Why Use linkeR?

### Before linkeR (Manual Approach)
```{r before-linker}
# Complex manual setup for each component pair
observeEvent(input$map_marker_click, {
  clicked_id <- input$map_marker_click$id
  # Find corresponding table row
  row_idx <- which(my_data()$id == clicked_id)
  # Update table selection
  dataTableProxy("my_table") %>% selectRows(row_idx)
  # Update map view
  leafletProxy("my_map") %>% setView(...)
  # Update any other components...
  # Repeat this for every component combination!
})

observeEvent(input$my_table_rows_selected, {
  # More boilerplate code...
  # Handle edge cases...
  # Ensure consistency...
})
```

### With linkeR (Simple Approach)
```{r after-linker}
# One line links everything!
link_plots(
  session,
  my_map = my_data,
  my_table = my_data,
  shared_id_column = "id"
)
```

### Key Benefits

| Manual Approach | linkeR Approach |
|-----------------|-----------------|
| 50+ lines of observer code | 1 function call |
| Easy to introduce bugs | Tested and reliable |
| Hard to maintain | Declarative and clear |
| Limited to 2 components | Unlimited components |
| No built-in customization | Rich customization options |

## Installation

```{r installation}
# Install from CRAN (when available)
install.packages("linkeR")

# Or install development version from GitHub
# install.packages("devtools")
devtools::install_github("JakeWags/linkeR")
```

## Requirements

For linking to work, your setup needs:

1. **Shared ID Column**: All datasets must have a common identifier column
2. **Matching Component IDs**: 
   - Leaflet: Use `layerId = ~your_id_column` 
   - DT: Row numbers automatically map to data rows
3. **Reactive Data**: Wrap your data in `reactive()`

```{r requirements-example}
# Good: Proper setup
my_data <- reactive({
  data.frame(
    id = 1:10,           # ← Shared ID column
    name = paste("Item", 1:10),
    lat = runif(10), lng = runif(10)
  )
})

output$my_map <- renderLeaflet({
  leaflet(my_data()) %>%
    addMarkers(layerId = ~id)  # ← layerId matches shared_id_column
})

link_plots(session, my_map = my_data, shared_id_column = "id")

# Bad: Missing layerId
output$my_map <- renderLeaflet({
  leaflet(my_data()) %>%
    addMarkers()  # ← No layerId = no linking!
})
```

## Supported Components

| Component | Status | Notes |
|-----------|--------|--------|
| Leaflet Maps | ✅ Full Support | Interactive maps with markers, circles, polygons |
| DT DataTables | ✅ Full Support | Sortable, filterable tables |
| Plotly Charts | 🔄 Partial | Requires manual event handling |
| Custom Components | 🔄 Partial | Any Shiny component with click events, Requires manual event handling |
| Base R Plots | 📋 Planned | Static plots with click detection |
| Mapbox | 📋 Planned | Alternative mapping solution |

## Contributing

Contributions are welcome and encouraged! Follow best practice for github contributions.

### Development Setup
```{r development}
# Clone and setup
git clone https://github.com/JakeWags/linkeR.git
cd linkeR

# Install dependencies
devtools::install_deps()

# Run tests
devtools::test()

# Check package
devtools::check()
```

## License

This project is licensed under the MIT License - see the [LICENSE](LICENSE) file for details.

---
