---
title: "Getting Started with linkeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with linkeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Set to FALSE since these are Shiny examples
)
```

```{r setup}
library(linkeR)
library(shiny)
library(leaflet)
library(DT)
```

# Introduction

The `linkeR` package provides a simple way to create bidirectional links between interactive components in Shiny applications. When users click on one component (like a map marker), other linked components (like tables) automatically update to show related information.

## Core Concept

The main idea is simple: **one click, multiple updates**. When you select an item in any linked component, all other linked components highlight or select the corresponding item.

## Basic Usage

The simplest way to link components is with the `link_plots()` function:

```{r basic-example}
library(shiny)
library(leaflet)
library(DT)
library(linkeR)

# Sample data with shared IDs
sample_data <- data.frame(
  id = 1:10,
  name = paste("Location", 1:10),
  latitude = runif(10, 40.7, 40.8),
  longitude = runif(10, -111.95, -111.85),
  value = runif(10, 100, 1000)
)

ui <- fluidPage(
  titlePanel("Basic linkeR Example"),
  
  fluidRow(
    column(6,
      h4("Map"),
      leafletOutput("my_map")
    ),
    column(6,
      h4("Table"),
      DTOutput("my_table")
    )
  )
)

server <- function(input, output, session) {
  
  # Create reactive data
  my_data <- reactive({ sample_data })
  
  # Render components
  output$my_map <- renderLeaflet({
    leaflet(my_data()) %>%
      addTiles() %>%
      addMarkers(
        lng = ~longitude,
        lat = ~latitude,
        layerId = ~id,  # Critical: this must match your shared_id_column
        popup = ~paste("Name:", name)
      )
  })
  
  output$my_table <- renderDT({
    datatable(
      my_data()[, c("name", "value")],
      selection = "single",
      rownames = FALSE
    )
  })
  
  # Set up linking - this is all you need!
  link_plots(
    session,
    my_map = my_data,      # component_name = reactive_data
    my_table = my_data,    # component_name = reactive_data
    shared_id_column = "id"  # column that links the components
  )
}

shinyApp(ui, server)
```

## Key Requirements

For linking to work, you need:

1. **Shared ID column**: All linked datasets must have a common identifier column
2. **Matching layerId/rowId**: Components must use the shared ID for selection
3. **Reactive data**: Data must be wrapped in `reactive()`

## What Happens When You Click?

- **Map marker click**: Table selects corresponding row
- **Table row click**: Map zooms to marker and shows popup
- **Consistent behavior**: Both interactions produce the same visual result

# Next Steps

- See `vignette("custom-behaviors")` for advanced customization
- See `vignette("multiple-components")` for complex linking scenarios
- See `vignette("troubleshooting")` for common issues and solutions