---
title: "Linking Multiple Components"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linking Multiple Components}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(linkeR)
library(shiny)
library(leaflet)
library(DT)
library(plotly)
```

# Linking Multiple Components

`linkeR` can link any number of interactive components together. This vignette shows how to create complex dashboards with maps, tables, charts, and other elements all responding to the same selections.

## Three-Component Example

Here's a dashboard with a map, table, and plotly chart all linked together:

```{r three-component}
ui <- fluidPage(
  titlePanel("Multi-Component Dashboard"),
  
  fluidRow(
    column(4,
      h4("Geographic View"),
      leafletOutput("map")
    ),
    column(4,
      h4("Data Table"),
      DTOutput("table")
    ),
    column(4,
      h4("Performance Chart"),
      plotlyOutput("chart")
    )
  ),
  
  fluidRow(
    column(12,
      h4("Selection Details"),
      verbatimTextOutput("selection_info")
    )
  )
)

server <- function(input, output, session) {
  
  # Sample business data
  business_data <- reactive({
    data.frame(
      business_id = paste0("BIZ_", 1:20),
      name = paste("Business", 1:20),
      latitude = runif(20, 40.7, 40.8),
      longitude = runif(20, -111.95, -111.85),
      revenue = runif(20, 100000, 1000000),
      employees = sample(10:100, 20, replace = TRUE),
      category = sample(c("Tech", "Retail", "Food"), 20, replace = TRUE)
    )
  })
  
  # Render map
  output$map <- renderLeaflet({
    leaflet(business_data()) %>%
      addTiles() %>%
      addCircleMarkers(
        lng = ~longitude,
        lat = ~latitude,
        layerId = ~business_id,
        radius = 5,
        popup = ~paste("Business:", name)
      )
  })
  
  # Render table
  output$table <- renderDT({
    data <- business_data()
    display_data <- data[, c("name", "category", "revenue", "employees")]
    
    datatable(
      display_data,
      selection = "single",
      rownames = FALSE
    ) %>%
      formatCurrency("revenue", currency = "$", digits = 0)
  })
  
  # Render chart
  output$chart <- renderPlotly({
    plot_ly(
      data = business_data(),
      x = ~employees,
      y = ~revenue,
      key = ~business_id,  # Critical: this enables plotly linking
      text = ~name,
      type = "scatter",
      mode = "markers",
      source = "business_chart"  # Source ID for plotly events
    ) %>%
      layout(
        title = "Revenue vs Employees",
        xaxis = list(title = "Employees"),
        yaxis = list(title = "Revenue")
      )
  })
  
  # Link the map and table (plotly requires special handling)
  registry <- link_plots(
    session,
    map = business_data,
    table = business_data,
    shared_id_column = "business_id"
  )
  
  # Handle plotly clicks separately
  observeEvent(event_data("plotly_click", source = "business_chart"), {
    clicked_data <- event_data("plotly_click", source = "business_chart")
    if (!is.null(clicked_data) && !is.null(clicked_data$key)) {
      # Update the registry selection
      registry$set_selection(clicked_data$key, "chart")
    }
  })
  
  # Display selection info
  output$selection_info <- renderText({
    selection <- registry$get_selection()
    if (!is.null(selection$selected_id)) {
      data <- business_data()
      selected <- data[data$business_id == selection$selected_id, ]
      if (nrow(selected) > 0) {
        paste0(
          "Selected: ", selected$name, "\n",
          "Source: ", selection$source, "\n",
          "Revenue: $", format(selected$revenue, big.mark = ","), "\n",
          "Employees: ", selected$employees
        )
      }
    } else {
      "No selection"
    }
  })
}
```

## Adding More Component Types

### Summary Tables

You can link summary tables that show aggregated information:

```{r summary-table}
# In your server function
output$category_summary <- renderDT({
  data <- business_data()
  
  summary_data <- data %>%
    group_by(category) %>%
    summarise(
      count = n(),
      avg_revenue = mean(revenue),
      total_employees = sum(employees),
      .groups = "drop"
    )
  
  datatable(
    summary_data,
    selection = "single",
    rownames = FALSE,
    options = list(pageLength = 5, searching = FALSE)
  ) %>%
    formatCurrency("avg_revenue", currency = "$", digits = 0)
})

# Link it by adding to your link_plots call
link_plots(
  session,
  map = business_data,
  table = business_data,
  summary = business_data,  # Same data, different view
  shared_id_column = "business_id"
)
```

### Time Series Charts

For dynamic charts that update based on selection:

```{r time-series}
# Reactive chart that updates with selection
output$monthly_trend <- renderPlotly({
  registry <- get_registry()  # However you store your registry
  selection <- registry$get_selection()
  
  if (!is.null(selection$selected_id)) {
    # Get monthly data for selected business
    selected_business <- business_data()[
      business_data()$business_id == selection$selected_id, 
    ]
    
    # Generate monthly data (example)
    monthly_data <- data.frame(
      month = 1:12,
      revenue = selected_business$revenue / 12 * runif(12, 0.8, 1.2)
    )
    
    plot_ly(
      data = monthly_data,
      x = ~month,
      y = ~revenue,
      type = "scatter",
      mode = "lines+markers"
    ) %>%
      layout(
        title = paste("Monthly Trend:", selected_business$name),
        xaxis = list(title = "Month"),
        yaxis = list(title = "Revenue")
      )
  } else {
    # Empty chart with instructions
    plot_ly() %>%
      add_annotations(
        text = "Select a business to view trends",
        x = 0.5, y = 0.5,
        xref = "paper", yref = "paper",
        showarrow = FALSE
      )
  }
})
```

## Managing Complex State

For complex dashboards, you might want centralized state management:

```{r state-management}
server <- function(input, output, session) {
  
  # Centralized selection state
  current_selection <- reactiveVal(NULL)
  
  # Business data
  business_data <- reactive({ /* your data */ })
  
  # Set up linking with global callback
  registry <- link_plots(
    session,
    map = business_data,
    table = business_data,
    shared_id_column = "business_id",
    
    # Update centralized state
    on_selection_change = function(selected_id, selected_data, source_id, session) {
      current_selection(selected_data)
      
      # Log for debugging
      cat("Selection changed:", selected_id, "from", source_id, "\n")
    }
  )
  
  # All outputs can react to current_selection()
  output$detail_panel <- renderUI({
    selected <- current_selection()
    if (!is.null(selected)) {
      # Rich detail panel
      fluidRow(
        column(6, 
          h4("Business Details"),
          p("Name:", selected$name),
          p("Category:", selected$category)
        ),
        column(6,
          h4("Performance"),
          p("Revenue:", scales::dollar(selected$revenue)),
          p("Employees:", selected$employees)
        )
      )
    } else {
      div("Select a business to view details")
    }
  })
  
  output$related_businesses <- renderDT({
    selected <- current_selection()
    if (!is.null(selected)) {
      # Show businesses in same category
      related <- business_data()[
        business_data()$category == selected$category & 
        business_data()$business_id != selected$business_id,
      ]
      
      datatable(
        related[, c("name", "revenue", "employees")],
        caption = paste("Other", selected$category, "businesses"),
        options = list(pageLength = 5)
      )
    }
  })
}
```

## Performance Tips

1. **Limit Updates**: Use `req()` to avoid unnecessary updates
2. **Debounce Fast Clicks**: Use `debounce()` for rapid selection changes
3. **Lazy Loading**: Only render expensive components when selected
4. **Efficient Data**: Keep linked datasets reasonably sized

## Troubleshooting Multiple Components

- **IDs must match**: Ensure all components use the same shared ID values
- **Reactive context**: All data must be in `reactive()` expressions
- **Observer conflicts**: Be careful with multiple `observeEvent()` on the same inputs
- **plotly requires special handling**: Use the registry's `set_selection()` method for plotly clicks

With these patterns, you can create rich, interconnected dashboards where every component enhances the user's understanding of the data.
